Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-12-05 paper/main.tex
rewrite amir part to match figure

Since this setup is void of any cell-cell interactions or cell division, the dynamics of the cell
can be described by a combination of growth and physical mechanics (see
subsection~\ref{subsection:mechanical-model-mechanics}).
We are thus left with 5 parameters:
Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (upright, light blue)
and intermediate (bent, dark blue) position of the rod using the optimized parameters.
The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (C-D, blue curves)
show that each parameter is identifiable from the given data.
However when it comes to the damping constant $\lambda$, its overall numerical value and steepness
of the curve compared to the results for the other parameters are much lower.
We can assess if the system could be described by a reduced model by setting $\lambda=0$, thus
effectively removing any damping.
The curves which are obtained in this way are displayed in orange and show that the system is also
fully identifiable in its reduced version.

It should be noted that the used images only contain snapshots of a fully-bent or fully-relaxed rod
without any intermediate steps.
This effectively limits us to effectively 2 datapoints and is likely the key factor that leads to
the previously described phenomenon.
The damping force is of dynamic nature, and takes no effect in the two steady states of the rod
being either fully bent or fully relaxed.
Gathering additional transient information would allow us to better estimate the damping parameter.

>>>>

Since this setup is void of any cell-cell interactions or cell division, the dynamics of the cell
can be described by a combination of growth and physical mechanics (see
subsection~\ref{subsection:mechanical-model-mechanics}).
We are thus left with 5 parameters:
Magnitude of the drag force $C$, growth rate $\alpha$, rigidity of the rod $\eta$, spring tension
$\gamma$ and damping constant $\lambda$ (see table~\ref{table:list-of-parameters}).
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (upright, light blue)
and intermediate (bent, dark blue) position of the rod using the optimized parameters.
The likelihood profiles displayed in figure~\ref{fig:amir-bending-simulation} (D-H, bright blue
curves) show that each parameter sits at minimum of the \texttt{profile-likelihood}.
However, Figure~\ref{fig:amir-bending-simulation} (E-G) exhibit multimodal minima which can not be
distinguished.
Furthermore, the damping constant $\lambda$ can not be distinguished from $\lambda=0$ within the
lowest confidence interval of $68\%$.
We can assess if the system could be described by a reduced model by setting $\lambda=0$, thus
effectively removing any damping.
The resulting curves (dark blue) still exhibit the multimodal minima but a further reduction of
fixing the spring tension $\gamma$ produces a (orange curves) fully identifiable system.
As we have explained earlier (Section~\ref{sec:computational-model}), $\gamma$ can be viewed as an
auxiliary parameter.

It should be noted that the used images only contain snapshots of a fully-bent or fully-relaxed rod
without any intermediate steps.
This effectively limits us to effectively 2 datapoints and is likely the key factor that leads to
the non-identifiability of the damping constant $\lambda$, since this force takes no effect in the
two steady states (fully bent or fully relaxed).
Gathering additional transient information would allow us to better estimate this parameter.
